NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PS lipids

### 2018-07-04 Manuscript/manuscriptPS.tex
Update authors and small thinbgs

\author{O. H. Samuli Ollila}
\email[]{samuli.ollila@helsinki.fi}
%\homepage[]{Your web page}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\affiliation{Institute of Biotechnology, University of Helsinki}

\author{NMRlipids collaboration \todo{Authorship query to be sent soon.}}
\affiliation{nmrlipids.blogspot.fi}

@@

Several classical molecular dynamics (MD) simulation studies are done
to understand PS headgroups, their influence on lipid bilayer properties and
interactions with ions \cite{cascales96,pandit02,mukhopadhyay04,pedersen06,vernier09,boettcher11,molina12,venable13,pan14,vangaveti14,melcrova16}.
However, the recent comparisons of PC lipid headgroup and glycerol backbone
C-H bond order parameters calculated from different simulation models
revealed that improvements in the current force fields are needed
to correctly reproduce the headgroup structure and ion binding to
lipid bilayers \cite{botan15,catte16,ollila16}. The ion binding affinity
to POPC bilayer was then improved by implicitly including the electronic
polarizability using the electronic continuum correction \cite{melcr18}.
Here, we collect the set of experimentally measured lipid headgroup and
glycerol backbone C-H bond order parameters, which can be used to
evaluate the quality of headgroup structure and the ion binding affinity
in MD simulations of lipid bilayers containing PS lipids. 
The available MD simulation models of PS are then compared against
the collected experimental data. The results pave the way for the
development of MD simulation force fields that correctly describe
PS lipid headgroup structure and its interactions with ions.
Such models are expected to be useful in elucidating the biological
role of PS and other lipid headgroups because glycerol backbone and
lipid headgroups behave similarly in model membranes and in
bacteria \cite{gally81,scherer87,seelig90}.

@@

    POPS/Na$^+$  & MacRog \cite{maciejewski14}  &0 & 0        & ? & ?? & 0  & ?  & ? & ?  & \cite{??} \todoi{Data to be added by Piggot} \\

@@

\begin{figure*}[]
  \centering
  \includegraphics[width=16.0cm]{../Figs/HGorderparametersPCvsPS.eps}
  \caption{\label{HGorderparametersPCvsPS}
    Changes of PC (left panel) and PS (right panel) headgroup order parameters
    from POPC:POPS mixtures with increasing amount of POPS.
    Experimental results of POPC are taken from Ref. \citenum{scherer87}
    (signs are determined as discussed in \cite{botan15,ollila16}).
    Experimental values for POPS in pure bilayer and in mixture are  
    measured in this work and in Ref. \citenum{roux90} at 298K, respectively.
    Since the experimental data of POPS in pure and diluted mixture come from
    different experimental sets (13C NMR in this work and 2H NMR from Ref. \citenum{roux90}),
    the experimental change of the order parameter is less accurate than in typical measurements
    where same technique is used in all conditions, see discussion about qualitative and quantitative 
    accuracy in Ref. \citenum{ollila16}.
    For POPC (left panel) the zero point of y-axis is set to the value of pure bilayer.
    For $\beta$-carbon of POPS (right panel, top) the zero point of y-axis is
    set to the value from POPC:POPS (5:1) mixture.
    For $\alpha$-carbon of POPS (right panel, bottom) the y-axis is tranferred
    with the same value for both order parameters such that the lower order
    parameter value from POPC:POPS (5:1) mixture is at zero
    to correctly illustrate the significant forking.
  }
  \todo{Simulation of CHARMM36 at 298K should be maybe rerun with Gromacs 5.} \\
  \todo{Simulation of pure POPC at 298K with Lipid14 would be useful for this plot (only at 303 K is available from NMRlipids I)} \\
  \todo{MacRog simulations of pure POPS with potassium counterions only would be useful for this and other plots.}
\end{figure*}

>>>>

\author{Tiago M. Ferreira}
\affiliation{Halle, Germany}
\author{Josef Melcr}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\author{O. H. Samuli Ollila  \todo{Authorlist is not yet complete}}
\email[]{samuli.ollila@helsinki.fi}
%\homepage[]{Your web page}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\affiliation{Institute of Biotechnology, University of Helsinki}
%\author{NMRlipids collaboration}
%\affiliation{nmrlipids.blogspot.fi}

@@

Several classical molecular dynamics (MD) simulation studies are done
to understand PS headgroups, their influence on lipid bilayer properties and
interactions with ions \cite{cascales96,pandit02,mukhopadhyay04,pedersen06,vernier09,boettcher11,molina12,jurkiewicz12,venable13,pan14,vangaveti14,melcrova16}.
However, the recent comparisons of PC lipid headgroup and glycerol backbone
C-H bond order parameters calculated from different simulation models
revealed that improvements in the current force fields are needed
to correctly reproduce the headgroup structure and ion binding to
lipid bilayers \cite{botan15,catte16,ollila16}. The ion binding affinity
to POPC bilayer was then improved by implicitly including the electronic
polarizability using the electronic continuum correction \cite{melcr18}.
Here, we collect the set of experimentally measured lipid headgroup and
glycerol backbone C-H bond order parameters, which can be used to
evaluate the quality of headgroup structure and the ion binding affinity
in MD simulations of lipid bilayers containing PS lipids. 
The available MD simulation models of PS are then compared against
the collected experimental data. The results pave the way for the
development of MD simulation force fields that correctly describe
PS lipid headgroup structure and its interactions with ions.
Such models are expected to be useful in elucidating the biological
role of PS and other lipid headgroups because glycerol backbone and
lipid headgroups behave similarly in model membranes and in
bacteria \cite{gally81,scherer87,seelig90}.

@@

    POPS/Na$^+$  & MacRog \cite{maciejewski14}  &0 & 0        & 128 & 4480 & 0  & 298  & 500 & 100  & \cite{macrogPOPS298Kcorrect} \\

@@

\begin{figure*}[]
  \centering
  \includegraphics[width=16.0cm]{../Figs/HGorderparametersPCvsPS.eps}
  \caption{\label{HGorderparametersPCvsPS}
    Changes of PC (left panel) and PS (right panel) headgroup order parameters
    from POPC:POPS mixtures with increasing amount of POPS.
    Experimental results of POPC are taken from Ref. \citenum{scherer87}
    (signs are determined as discussed in \cite{botan15,ollila16}).
    Experimental values for POPS in pure bilayer and in mixture are  
    measured in this work and in Ref. \citenum{roux90} at 298K, respectively.
    Since the experimental data of POPS in pure and diluted mixture come from
    different experimental sets (13C NMR in this work and 2H NMR from Ref. \citenum{roux90}),
    the experimental change of the order parameter is less accurate than in typical measurements
    where same technique is used in all conditions, see discussion about qualitative and quantitative 
    accuracy in Ref. \citenum{ollila16}.
    For POPC (left panel) the zero point of y-axis is set to the value of pure bilayer.
    For $\beta$-carbon of POPS (right panel, top) the zero point of y-axis is
    set to the value from POPC:POPS (5:1) mixture.
    For $\alpha$-carbon of POPS (right panel, bottom) the y-axis is tranferred
    with the same value for both order parameters such that the lower order
    parameter value from POPC:POPS (5:1) mixture is at zero
    to correctly illustrate the significant forking.
  }
  \todo{Simulation of CHARMM36 at 298K should be maybe rerun with Gromacs 5.} \\
  \todo{Simulation of pure POPC at 298K with Lipid14 would be useful for this plot (only at 303 K is available from NMRlipids I)} \\
  \todo{MacRog simulations of pure POPS with potassium counterions only would be useful for this and other plots.} \\
  \todo{The data for Gromos-CKP by Tom Piggot: http://nmrlipids.blogspot.com/2017/03/nmrlipids-iv-headgroup-glycerol.html?showComment=1528213129976\#c8319083196182661431
    should be added into the plot.}
\end{figure*}
